\section{Square loss exemplar machine}\label{lsesvm}
In this section, we revisit the exemplar SVM model as presented by \cite{Efros11} and introduce its square loss version.
\subsection{Exemplar SVMs} \label{esvm}
Given features in $\RR^d$ at training time, one positive example $x_0$ in $_RR^d$ and a set of negative examples $X = [x_1, x_2,...,x_n]$, in $\RR^{d\times n}$each column of $X$ representing one example by a vector in $\RR^d$. 
Any feature vector representation can be used.
We are also given a loss function $l:\{-1, 1\}\times \RR\rightarrow\RR^+$. Learning an E-SVM from these examples amounts to minimizing the function
\begin{equation}
J(\omega, \nu) = \theta \ l(1, \omega^Tx_0+\nu) +\dfrac{1}{n}\sumi l(-1, \omega^T x_i+\nu)+\dfrac{\lambda}{2}|\!|\omega|\!|^2, \label{eq:first}
\end{equation}
w.r.t. $\omega$ in $\RR^d$ and $\nu$ in $\RR$.

In Equation (\ref{eq:first}), $\lambda$ and $\theta$ are respectively a regularization parameter on $\omega$ and a positive scalar adjusting the weight of the positive exemplar.
%\footnote{We could also acknowledge a parameter $\theta$ other than $\frac{1}{n}$ as a regularization to the error of each negative example. But this parameters seems to be less important to cross-validate. Also, setting $\theta=\frac{1}{n}$ simplify Equation (\ref{omega:solution}) and allows the use of Woodbury identity.} 

The  exemplar SVM of $x_0$ with respect to $\NNN$ is the classifier $\omega^\star$ that minimizes the function $J$:
\begin{equation}
\omega^\star(x_0, X), \nu^\star(x_0, X) = \underset{\omega\in\RR^d, \ \nu\in\RR}{argmin} \ J(\omega, \nu). \label{omega:first}
\end{equation}
To shorten the notation, we refer to these solutions as $\omega^\star$ and $\nu^\star$ when their arguments are implicit.\\
Applications of E-SVM use the hinge loss function, which guarantees that $J$ is a convex function. The solution of Equation (\ref{omega:first}) can be thus found by stochastic gradient descent.

\subsection{Square loss function}\label{SLEM}
Now, let us study the same learning problem for the square-loss function $l(y,\hat{y}) = \frac{1}{2}(y-\hat{y})^2$. As for the hinge loss, Equation (\ref{eq:first}) is a convex problem. 
However, differently from the hinge loss, Equation (\ref{eq:first}) is a ridgr regression problem, whose solution is expressed in a close form:

\begin{align}
\begin{cases}
\vspace{3 mm}
\nu^\star &= \dfrac{\theta-1}{\theta+1}-\dfrac{1}{\theta+1}(\theta x_0+\mu)^T\omega^\star,\\
\vspace{3 mm}
\omega^\star &= \dfrac{2\theta}{\theta+1}U^{-1}(x_0-\mu), \ where \\
\end{cases}
\label{omega:solution}
\end{align}

\begin{align}
where \ \mu = \frac{1}{n}\sum_{i=1}^n x_i, \ \ U = \dfrac{1}{n}XX^T-\mu\mu^T+\dfrac{\theta}{\theta+1}(x_0-\mu)(x_0-\mu)^T+\lambda\textbf{Id}_d.
\end{align}
Equation (\ref{omega:solution}) shows how to solve (\ref{eq:first}) in a closed form to obtain the SLEM vector $\omega^\star(x_0,X)$ of $x_0$. 
Replacing the hinge loss by the square loss offers a more compact solution, but not necessarily more efficient.

\subsection{Woodbury identity and matching scores}
Let us define $A = \frac{1}{n}XX^T-\mu\mu^T +\lambda\textbf{Id}_d$ and assume $A^{-1}$ known. 
%$A$ is the covariance matrix of the negative samples $X$\footnote{added of a small term proportional to the identity matrix to insure $A$ is positive-definite.}. Let us assume its inverse $A^{-1}$ is known. 
We know that $U = A + \dfrac{\theta}{\theta+1}\delta\delta^T$, where $\delta=x_0-\mu$. The Woodbury identity \cite{woodbury} gives us
\begin{equation}
U^{-1} = A^{-1} -\dfrac{\theta}{\theta\delta^TA^{-1}\delta+ \theta+1}A^{-1}\delta^T\delta A^{-1}. \label{invU}
\end{equation}

Replacing (\ref{invU}) in the original equation
\begin{align}
\omega^\star &= \dfrac{2\theta}{\theta+1}U^{-1}\delta \\
&= A^{-1}\delta - \dfrac{\theta}{\theta\delta^TA^{-1}\delta+ \theta+1} A^{-1}\delta (\delta^TA^{-1}\delta)\\
&= \dfrac{2\theta}{\theta\delta^TA^{-1}\delta+ \theta+1} A^{-1}\delta.\label{Wood:omega}
\end{align}

The first observation we make from Equation (\ref{Wood:omega}) is the positive sample weight $\theta$ does not influence the direction of the optimal vector $\omega^\star$, only its norm. This means that according to the matching score criteria defined by \cite{ZePe15}, $\theta$ does not influence the matching score of the SLEM vectors of two different images.

Indeed, if $\omega$ and $\omega'$ are the $d$-dimensional SLEM vectors of $x$ and $x'$, respectively, we can denote $s$ the matching score scalar function defined in $\RR^d\times \RR^d$, which is given by
\begin{equation}
s(\omega, \omega') = \dfrac{1}{|\omega||\omega'|}\langle\omega,\omega'\rangle = (x-\mu)^TA^{-2}(x'-\mu).\label{match:score}
\end{equation}

Equation (\ref{match:score}) is independent from the value of $\theta$. This means that the weight of the positive sample can be fixed at any positive value without changing matching scores. This sets the SLEM appart from E-SVM, where in this parameter had to be cross validated \cite{Efros11, ZePe15}.
\subsection{LDA and SLEM}

Let us now reanalyse the SLEM problem and suppose that we have multiple positive samples. It can be shown that in this case, the corresponding linear classifier of Equation (\ref{eq:first}) for the square-loss is also given by
Equation (\ref{omega:solution}), where $x_0$ denotes this time the center of mass
of the positive samples, {\em if} these samples have the {\em same} covariance matrix $\Sigma$ of the negative samples $X$.
 
This equal-covariance assumption is of course quite restrictive, and
probably unrealistic in general. It is interesting to note, however,
that this is exactly the assumption made by linear discriminant
analysis. As shown in~\cite{Hastie2009} for example, LDA can be seen as a
(non-regularized) linear classifier
with decision function $w'\cdot z+ b'$, where $z$ is a sample in
$\RR^d$, and

\begin{equation}
\left\{\begin{array}{l}
\displaystyle w'=\Sigma^{-1}(x_0-\mu),\\
\displaystyle b'=-\frac{1}{2}(x_0+\mu)^T w',
\end{array}\right.
\label{eq:lda}
\end{equation}
 
Note that when $\lambda=0$ (no regularization),
\begin{align}
\Sigma w & = Uw -\frac{1}{2}[(x_0-\mu)^T w] (x_0-\mu) =
\left[1-\frac{1}{2}(x_0-\mu)^T w\right](x_0-\mu) \\
&=\left[1-\frac{1}{2}(x_0-\mu)^T w\right]\Sigma w'
\end{align}
thus $w$ and $w'$ have the same direction. In other words, a
SLEM is a generalized version of LDA when the
regularizer parameter $\lambda $ is zero.
This observation has been equally made by \cite{Koba15}.

Many interesting properties of LDA has been rediscovered classification tasks \cite{GMPD12,HMR12} and, more recently, for image retrieval with non-embedded image encoders \cite{babenko15}.
